---
keep_md: yes
pagetitle: EDA - Real State Sales
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
```

## Analysis of Real State Sales

We want to look at the impact of the second avenue line by looking at how the real estate sales prices have changed in the areas surrounding the subway lines. Similar to looking at building permits, we are looking at how the prices have changed by block. 

To reiterate, we're using the NYC Department of Finance data on the real estate transactions in the zip codes surrounding the Upper East Side. We've narrowed down are real estate transactions to properties that are in tax two, and for sales prices between 100K and 5M. Here, we're focusing on phase 1 and 2, since phase 3 and 4 are not built yet. 


```{r}
library(tidyverse)
library(ggplot2)
library("readxl")
library(reshape)
library(dplyr)
library("RSQLite")

```

```{r}
data_ues_1718  <- read.csv(file="data/all_ues_sales_geocode_2017-2018.csv", header=TRUE, sep=",")
colnames(data_ues_1718)[colnames(data_ues_1718) == "SALE.PRICE"] <- 'price'
colnames(data_ues_1718)[colnames(data_ues_1718) == "YEAR"] <- 'year'
colnames(data_ues_1718)[colnames(data_ues_1718) == "ZIP.CODE"] <- 'zipcode'
colnames(data_ues_1718)[colnames(data_ues_1718) == "TAX.CLASS.AT.PRESENT"] <- 'TaxClass'
data_ues_1718 <- data_ues_1718[data_ues_1718$price > 100000,]
data_ues_1718 <- data_ues_1718[data_ues_1718$price < 5000000,]
```


```{r}
sqlite<- dbDriver("SQLite")
realestate <- dbConnect(sqlite,"/Users/sharontsao/Desktop/NYC_RealEstate_Sales.sqlite")

#dbListTables(realestate)

#data_14 = dbGetQuery(realestate,'select * from yr_2014')

data_ues = dbGetQuery(realestate,'

                  select year, nbhd, block, lot, address, zip, sale_date, price, long, lat
                      from yr_2016 where zip in (10065,10021, 10075, 10028 ,10029, 10128) 
                      AND usable = "True" 
                      AND tax_cls_s = 2
                      AND price > 100000
                      AND price < 5000000
                      
            UNION
                      select year, nbhd, block, lot, address, zip, sale_date, price, long, lat
                      from yr_2015 where zip in (10065,10021, 10075, 10028 ,10029, 10128) 
                      AND usable = "True" 
                      AND tax_cls_s = 2
                      AND price > 100000
                      AND price < 5000000

            UNION

                      select year, nbhd, block, lot, address, zip, sale_date, price, long, lat
                      from yr_2014 where zip in (10065,10021, 10075, 10028 ,10029, 10128) 
                      AND usable = "Y" 
                      AND tax_cls_s = 2
                      AND price > 100000
                      AND price < 5000000');


write.csv(data_ues, 'ues_data_14-16.csv')
```

```{r}
# Left from 2nd Avenue Line

data_phase1_left_1718 <- data_ues_1718 %>%
  mutate(Distance = ifelse(data_ues_1718$BLOCK %in% c(1424:1427, 1529:1532,1539:1541,1423,1428:1433,
                                        1525:1528,1533:1538,1646),
                                 '1 block left from 2nd Ave line',
                          ifelse(data_ues_1718$BLOCK %in% c(1404:1407,1512:1515,1522:1524,1403,1408:1413,
                                              1508:1511,1516:1521,1624),
                                 '2-3 blocks left from 2nd Ave line',
                          ifelse(data_ues_1718$BLOCK %in% c(1384:1387,1495:1498,1505:1507,1383,1388:1393,
                                              1491:1494,1499:1504,1602),
                                 '4-5 blocks left from 2nd Ave line',
                                 'other'))))



data_phase1_left <- data_ues %>%
  mutate(Distance = ifelse(data_ues$block %in% c(1424:1427, 1529:1532,1539:1541,1423,1428:1433,
                                        1525:1528,1533:1538,1646),
                                 '1 block left from 2nd Ave line',
                          ifelse(data_ues$block %in% c(1404:1407,1512:1515,1522:1524,1403,1408:1413,
                                              1508:1511,1516:1521,1624),
                                 '2-3 blocks left from 2nd Ave line',
                          ifelse(data_ues$block %in% c(1384:1387,1495:1498,1505:1507,1383,1388:1393,
                                              1491:1494,1499:1504,1602),
                                 '4-5 blocks left from 2nd Ave line',
                                 'other'))))
```
### Analysis of Real Estate Sales Prices for Phase 1

To see how the subway has affected the real estate prices, we have added a distance column, which measure the number of blocks away from the subway, from the right/left side of the new Second Avenue line. 

The reason why we are differentiating between blocks on the right and the left side is that the left side of the Second Avenue line has the current 4/5/6 line present, and that real estate prices there have probably already been saturated because of the 4/5/6 line. We conjectured that the right side of the Second Avenue line may have a bigger impact. 

### Properties on the left side of the 2nd Ave Line - Phase 1

Here, we visualize the data by summarizing sales date and distance, and scaling the data to 2014 as an index.

Given that the property values were stemmed to the same index in 2014, the real estate prices that are closest to the subway have grown dramatically compared to areas that are farther away. The Q line was opened on 1/1/2017, and we can see that there was a price inflection in 2017. 


```{r}
# Left from 2nd Avenue Line
#First, summarize by Filing.Date year and Distance:
library(lubridate) #dates library

data_phase1_left_edited_1416 <- data_phase1_left %>%
  group_by(year, Distance) %>%
  summarise(Mean = mean(price, na.rm=TRUE)) %>%
  arrange(year) #arranges from lowest to highest value

data_phase1_left_edited_1416$year <- as.numeric(data_phase1_left_edited_1416$year)

data_phase1_left_edited_1718 <- data_phase1_left_1718 %>%
  group_by(year, Distance) %>%
  summarise(Mean = mean(price, na.rm=TRUE)) %>%
  arrange(year) 

year_factor <- factor(data_phase1_left_edited_1416$year)

data_phase1_left_edited = rbind(data_phase1_left_edited_1416, data_phase1_left_edited_1718)

#Now, scale the data and create an index to visualize variations over time, then plot:
data_phase1_left_edited <- data_phase1_left_edited %>% group_by(Distance) %>%
  filter(year >= 1990) %>%
  filter(Distance != "other") %>%
  mutate(index = 100*Mean/Mean[1]) %>% ungroup()

year_factor <- factor(data_phase1_left_edited$year)

ggplot(data_phase1_left_edited, aes(x= year_factor, y = index, group = Distance, color = Distance)) + geom_line(stat='identity', lwd = 0.75) +
    ggtitle("Average Real Estate Sales vs Distance to 2nd Ave (Phase1-left)") +
    labs (x = "Year", y = "Real Estate Price Index") +
    theme_grey(16) +
    theme(legend.title = element_blank())
```

**Conclusions:** From the graph above, it seems that the prices are correlated by the availability of the new subway stations. 

  From the graph above, blocks one block left of the subway line have grew up to an index of ~115 since the Q line opening, blocks two/three blocks left of the subway line have stayed around the same, and blocks that are four/five blocks away have dropped to below 90 of index recently (2018). However, prices dipped for blocks that are one block away on 2017 while others grew a little bit. It's hard to attribute what exactly is causing the sharp price change. 

```{r}
# Right from 2nd Avenue Line

data_phase1_right_1718 <- data_ues_1718 %>%
  mutate(Distance = ifelse(data_ues_1718$BLOCK %in% c(1444:1447,1546:1549,1556:1558,1443,1448:1453,
                                        1542:1545,1550:1555,1668),
                                 '1 block right from 2nd Ave Line',
                          ifelse(data_ues_1718$BLOCK %in% c(1464:1467,1563:1566,1573,1463,1468:1473,
                                              1559:1562,1567:1571,1690),
                                 '2 blocks right from line',
                          ifelse(data_ues_1718$BLOCK %in% c(1482:1484,1580:1583,1480,1485:1490,1576:1579,
                                              1584:1587,1590,1592,1589),
                                 '3-4 blocks right from 2nd Ave Line',
                                 'other'))))


data_phase1_right <- data_ues %>%
  mutate(Distance = ifelse(data_ues$block %in% c(1444:1447,1546:1549,1556:1558,1443,1448:1453,
                                        1542:1545,1550:1555,1668),
                                 '1 block right from 2nd Ave Line',
                          ifelse(data_ues$block %in% c(1464:1467,1563:1566,1573,1463,1468:1473,
                                              1559:1562,1567:1571,1690),
                                 '2 blocks right from line',
                          ifelse(data_ues$block %in% c(1482:1484,1580:1583,1480,1485:1490,1576:1579,
                                              1584:1587,1590,1592,1589),
                                 '3-4 blocks right from 2nd Ave Line',
                                 'other'))))

```

### Properties on the right side of the 2nd Ave Line - Phase 1

Here, we visualize the data by summarizing sales date and distance, and scaling the data to 2014 as an index.

Given that the property values were stemmed to the same index in 2014, the real estate prices that are one block away from the subway have grown dramatically while blocks that are two blocks away have dipped, especially since 2017. Blocks that are three/four away have grew with the one block.

```{r}
# Right from 2nd Avenue Line
#First, summarize by Filing.Date year and Distance:
library(lubridate) #dates library


data_phase1_right_edited1416 <- data_phase1_right %>%
  group_by(year, Distance) %>%
  summarise(Mean = mean(price, na.rm=TRUE)) %>%
  arrange(year) #arranges from lowest to highest value

data_phase1_right_edited1416$year <- as.numeric(data_phase1_right_edited1416$year)

data_phase1_right_edited1718 <- data_phase1_right_1718 %>%
  group_by(year, Distance) %>%
  summarise(Mean = mean(price, na.rm=TRUE)) %>%
  arrange(year) 

year_factor <- factor(data_phase1_right_edited1416$year)

data_phase1_right_edited = rbind(data_phase1_right_edited1416, data_phase1_right_edited1718)


#Now, scale the data and create an index to visualize variations over time, then plot:
data_phase1_right_edited <- data_phase1_right_edited %>% group_by(Distance) %>%
  filter(year >= 1990) %>%
  filter(Distance != "other") %>%
  mutate(index = 100*Mean/Mean[1]) %>% ungroup()

year_factor <- factor(data_phase1_right_edited$year)

#Plot time series:
ggplot(data_phase1_right_edited, aes(x= year_factor, y = index, group = Distance, color = Distance)) + geom_line(stat='identity', lwd = 0.75) +
    ggtitle("Average Real Estate Sales vs Distance to 2nd Ave (Phase1-left)") +
    labs (x = "Year", y = "Indexed of Average Real Estate Sale") +
    theme_grey(16) +
    theme(legend.title = element_blank())

```

**Conclusions:** From the graph above, it seems like the new subway line has some effects on the real estate prices, but there probably is a bigger driver that's keep the three/four blocks right of the Second Avenue line in sync with the One block. 

From the Building Permits observation, we think that locations that are three/four blocks right from the Second Avenue line may always be a popular area because buildings there have a water view. 

#####**Analysis of Market Values for Phase 2**



```{r}

data_phase2_1718 <- data_ues_1718 %>%
  mutate(Distance = ifelse(data_ues_1718$BLOCK %in% c(1654:1656, 1665:1667,1788:1791,1647:1652,1659:1662,
                                        1783:1787), '1 block left from phase 2 Line',
                          ifelse(data_ues_1718$BLOCK %in% c(1676:1678, 1687:1689, 1801:1803,1669:1675,
                                              1680:1684,1795:1797),'1 block right from phase 2 Line',
                          ifelse(data_ues_1718$BLOCK %in% c(1625:1640,1643:1645,1767:1776),
                                 '2-3 blocks left from phase 2 Line',
                          ifelse(data_ues_1718$BLOCK %in% c(1691,1693:1696,1699:1705,1708:1711,1713,1715:1716,                                    1806:1808,1815,1816,1811,1813),
                                 '2-3 blocks right from phase 2 Line',
                                 'other')))))

data_phase2_1416 <- data_ues %>%
  mutate(Distance = ifelse(data_ues$block %in% c(1654:1656, 1665:1667,1788:1791,1647:1652,1659:1662,
                                        1783:1787), '1 block left from phase 2 Line',
                          ifelse(data_ues$block %in% c(1676:1678, 1687:1689, 1801:1803,1669:1675,
                                              1680:1684,1795:1797),'1 block right from phase 2 Line',
                          ifelse(data_ues$block %in% c(1625:1640,1643:1645,1767:1776),
                                 '2-3 blocks left from phase 2 Line',
                          ifelse(data_ues$block %in% c(1691,1693:1696,1699:1705,1708:1711,1713,1715:1716,                                    1806:1808,1815,1816,1811,1813),
                                 '2-3 blocks right from phase 2 Line',
                                 'other')))))

```




```{r}
#First, summarize by Year and Distance:
data_phase2_1416_edited <- data_phase2_1416 %>% 
  group_by(year, Distance) %>% 
  summarise(Mean = mean(price, na.rm=TRUE)) %>%
  arrange(year)

data_phase2_1416_edited$year <- as.numeric(data_phase2_1416_edited$year)

data_phase2_1718_edited1718 <- data_phase2_1718 %>%
  group_by(year, Distance) %>%
  summarise(Mean = mean(price, na.rm=TRUE)) %>%
  arrange(year)

year_factor <- factor(data_phase2_1416_edited$year)

data_phase2 = rbind(data_phase2_1718_edited1718, data_phase2_1416_edited)

#Now, scale the data and create an index to visualize variations over time, then plot:
data_phase2 <- data_phase2 %>% group_by(Distance) %>%
  filter(Distance != "other") %>%
  mutate(index = 100*Mean/Mean[1]) %>% ungroup()

year_factor <- factor(data_phase2$year)

#Plot time series:
ggplot(data_phase2, aes(x= year_factor, y = index, group = Distance, color = Distance)) + 
  geom_line(stat='identity', lwd = .75) +
    ggtitle("Average Real Estate Sales Vs Distance to 2nd Ave (Phase 2)") +
    labs (x = "Year", y = "Indexed of Average Real Estate Sale") +
    theme_grey(16) +
    theme(legend.title = element_blank()) 


```
