---
keep_md: yes
pagetitle: Interactive Component
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
```
See: [Our Exploratory Data Analysis (EDA) Journey](EDA-Building_Permits.html "Our EDA Journey")


## Interactive Component

One of the analysis we performed, was to find out if the NYC 2nd Ave Line subway project has affected the construction of new buildings (and remodeling of existing ones) in it's area of influence.

###**Did the NYC 2nd Ave Subway (yellow line), increase the construction/remodeling of buildings in the area?**

```{r, include=FALSE}
library(tidyverse)
permits_upper_east <- read.csv("data/DOB_Permit_Issuance.csv")
#Source of Data: https://data.cityofnewyork.us/Housing-Development/DOB-Permit-Issuance/ipu4-2q9a
#the data in the link above was previously pre-processed using the following code:
# permits <- read.csv("DOB_Permit_Issuance.csv")
# permits_upper_east <- filter(permits, BOROUGH=="MANHATTAN" & Zip.Code %in% c(10065, 10021, 10075, 10028, 10128, 10029, 10035))
# 
# #Select relevant columns of our interest:
# permits_upper_east <- select(permits_upper_east, BOROUGH,House..,Street.Name,Job.Type,Block,Lot, Community.Board, Zip.Code,Bldg.Type,Residential,Work.Type,Permit.Type,Permit.Subtype,Filing.Date,Issuance.Date,Expiration.Date,Job.Start.Date,Owner.s.Business.Type,LATITUDE,LONGITUDE,COUNCIL_DISTRICT,NTA_NAME)
# head(permits_upper_east)
# write.table(permits_upper_east, file="DOB_Permit_Issuance-up_east_harlem.csv",sep=",",row.names=F)

permits_upper_east$Filing.Date = as.Date(permits_upper_east$Filing.Date, format = "%m/%d/%Y")
permits_upper_east$Issuance.Date = as.Date(permits_upper_east$Issuance.Date, format = "%m/%d/%Y")
permits_upper_east$Job.Start.Date = as.Date(permits_upper_east$Job.Start.Date, format = "%m/%d/%Y")
permits_upper_east$Block = as.numeric(permits_upper_east$Block)
#head(permits_upper_east)
```


```{r fig.height=10, fig.width=10, include=FALSE}
#install.packages('rgdal') #read in general GIS package
#install.packages("leaflet") #interactive mapping
library(rgdal)
library(leaflet)

#Read NY Balocks and lots Shapefile 
NYC_Map <- readOGR("shp/nyu_MapPluto", "nyu_2451_34522") #read main Shapefile (takes long time)
#head(NYC_Map@data)
#Data from: https://geo.nyu.edu/catalog/nyu_2451_34522

#Read NYC Subway Station Lines, Stations and Entrances Shapefiles:
subway_lines <- readOGR("shp/Subway Lines", "geo_export_37a43bfe-d16c-4615-a1bb-bb5ea38465cb")
#head(subway_lines@data)
#Data from: https://data.cityofnewyork.us/Transportation/Subway-Lines/3qz8-muuu/data

subway_stations <- readOGR("shp/Subway Stations", "geo_export_c6d66a0c-2af5-4542-a222-be6844df9051")
#head(subway_stations@data)
#Data from: https://data.cityofnewyork.us/Transportation/Subway-Stations/arq3-7z49
```


```{r, include=FALSE}
library(lubridate) #dates library
all_permits <- permits_upper_east %>%
  group_by(Year = year(Filing.Date), Block) %>% #create variable "Year"
  summarise(Total = n()) %>%                    #total permits filed by block per year
  arrange(Year)                                 #arrange from lowest to highest year
#head(all_permits)
#write.table(all_permits,file="all_permits.csv",sep=",",row.names=F)

#Variation index by Block by year: 
permits_var <- all_permits %>%
  filter(Year >= 2001 & Year <= 2015) %>%
  group_by(Block) %>%
  mutate(index = Total/Total[1]-1) %>% ungroup()
#head(permits_var)

permits_var <- permits_var %>%
mutate(Distance = ifelse(Block %in% c(1424:1427, 1529:1532,1539:1541,1423,1428:1433,
                                        1525:1528,1533:1538,1646,1444:1447,1546:1549,1556:1558,1443,
                                      1448:1453,1542:1545,1550:1555,1668), 
                                  '1 block',
                          ifelse(Block %in% c(1404:1407,1512:1515,1522:1524,1403,1408:1413,
                                              1508:1511,1516:1521,1624),
                                 '2-3 blocks left',
                          ifelse(Block %in% c(1384:1387,1495:1498,1505:1507,1383,1388:1393,
                                              1491:1494,1499:1504,1602),
                                 '4-5 blocks left',
                          ifelse(Block %in% c(1464:1467,1563:1566,1573,1463,1468:1473,
                                              1559:1562,1567:1571,1690),
                                 '2 blocks right',
                          ifelse(Block %in% c(1482:1484,1580:1583,1480,1485:1490,1576:1579,
                                              1584:1587,1590,1592,1589),
                                 '3-4 blocks right',
                                 'other'))))))

phase1 <- c(1424:1427,1529:1532,1539:1541,1423,1428:1433,1525:1528,1533:1538,1646,1444:1447,1546:1549,1556:1558,1443,1448:1453,1542:1545,1550:1555,1668,1404:1407,1512:1515,1522:1524,1403,1408:1413,1508:1511,1516:1521,1624,1384:1387,1495:1498,1505:1507,1383,1388:1393,1491:1494,1499:1504,1602,1464:1467,1563:1566,1573,1463,1468:1473,1559:1562,1567:1571,1690,1482:1484,1580:1583,1480,1485:1490,1576:1579,1584:1587,1590,1592,1589)

write.table(permits_var,file="permits_var.csv",sep=",",row.names=F)

#Average Var Index per Distance:
permits_var[is.na(permits_var)] <- 0 #replace NA's by zero
permits_var <- permits_var %>%
  filter(Block != 1111) %>% #this is central park!
  group_by(Year,Distance) %>%
  mutate(avg.index = mean(index, na.rm = TRUE))

write.table(permits_var,file="permits_avg_var.csv",sep=",",row.names=F)
#unique(permits_var$Distance)
#head(permits_var)
```


```{r, include=FALSE}
#Before merging, lets spread permits_var so each year is a column containing the avg.index: 
permits_var <- permits_var %>%
  select(-contains("Total"), -starts_with("index")) 
permits_var <- spread(permits_var, Year, avg.index)
#head(permits_var)
colnames(permits_var)[3:17] <- paste("Y", colnames(permits_var)[3:17], sep = "_")
#colnames(permits_var)
write.table(permits_var,file="spread_permits_avg_var.csv",sep=",",row.names=F)
```


```{r fig.height=5, fig.width=7, echo=FALSE}
#Filter shapefile by zipcodes of interest:
NYC_Map2 <- NYC_Map[NYC_Map@data$ZipCode %in% c(10065,10021,10075,10028,10128,10029,10035) ,]
#Remove Central Park:
NYC_Map2 <- NYC_Map2[NYC_Map2@data$Block != 1111,]
#Select Phase 1 Blocks only: 
NYC_Map2 <- NYC_Map2[NYC_Map2@data$Block %in% phase1 ,]
#head(NYC_Map2@data)

# #Before merging, lets filter permits_var index by year 2015 (so we get cummulative variation index): 
# permits_var <- filter(permits_var, Year == 2015)

#merge both datasets and replace NA's with zero:
NYC_Map2 <- merge(NYC_Map2, permits_var, by="Block")
NYC_Map2@data[is.na(NYC_Map2@data)] <- 0
NYC_Map2@data$Distance[NYC_Map2@data$Distance==0] <- "other" #replace 0 Distance with "other"
NYC_Map2@data$avg.index[NYC_Map2@data$Distance=="other"] <- 0
#head(NYC_Map2@data)

write.table(NYC_Map2@data,file="NYC_Map2_data.csv",sep=",",row.names=F)

#set color scale to be used when mapping variation index for building permits:
palette <- colorNumeric(palette = "Blues", domain = NYC_Map2$Y_2015) #continuous range

library(scales)
library(htmltools)

#Set interactive labels for 2001-2005 Layer: 
block_labels2005 <- sprintf("<strong>%s from 2nd Ave Line:</strong> %s avg. increase<br/>in Bldg Permits (2001-2005)<br/>Zip Code: %s, Block: %s, Lot: %s<br/>Address: %s",NYC_Map2$Distance,
                            ifelse(NYC_Map2@data$Distance=='other',"NA", percent(NYC_Map2$Y_2005)),
                            NYC_Map2$ZipCode, NYC_Map2$Block, NYC_Map2$Lot, NYC_Map2$Address) %>%
  lapply(htmltools::HTML)

#Set interactive labels for 2001-2010 Layer: 
block_labels2010 <- sprintf("<strong>%s from 2nd Ave Line:</strong> %s avg. increase<br/>in Bldg Permits (2001-2010)<br/>Zip Code: %s, Block: %s, Lot: %s<br/>Address: %s",NYC_Map2$Distance,
                            ifelse(NYC_Map2@data$Distance=='other',"NA", percent(NYC_Map2$Y_2010)),
                            NYC_Map2$ZipCode, NYC_Map2$Block, NYC_Map2$Lot, NYC_Map2$Address) %>%
  lapply(htmltools::HTML)

#Set interactive labels for 2001-2015 Layer Shapefile: 
block_labels2015 <- sprintf("<strong>%s from 2nd Ave Line:</strong> %s avg. increase<br/>in Bldg Permits (2001-2015)<br/>Zip Code: %s, Block: %s, Lot: %s<br/>Address: %s",NYC_Map2$Distance,
                            ifelse(NYC_Map2@data$Distance=='other',"NA", percent(NYC_Map2$Y_2015)),
                            NYC_Map2$ZipCode, NYC_Map2$Block, NYC_Map2$Lot, NYC_Map2$Address) %>%
  lapply(htmltools::HTML)


#Set interactive labels for Subway Stations: 
stations_labels <- sprintf("<strong>Station:</strong> %s<br/><strong>Line:</strong> %s",
                           subway_stations$name, subway_stations$line) %>%
  lapply(htmltools::HTML)

#Set interactive labels for Subway Lines: 
line_labels <- ifelse(subway_lines$name=="4-5-6", "<strong>Pre-existing 4th Avenue Line</strong>",
                          ifelse(subway_lines$name=="Q","<strong>2nd Ave Line</strong>", "Other lines")) %>% lapply(htmltools::HTML)

myNYCmap <- leaflet(options = leafletOptions(minZoom = 9, maxZoom = 18)) %>% #restringe/limit Zoom
  #addTiles() %>%  # Add default OpenStreetMap base map
  addProviderTiles(providers$CartoDB.Positron) %>% #this base map looks better (see http://leaflet-extras.github.io/leaflet-providers/preview/index.html for other base maps)
  
  setView(lng = -73.954477, lat = 40.776522, zoom = 14.8) %>% #initial center point and zoom
  
  #addMarkers(lng=-73.954477, lat=40.776522, label="I can write some pop-up where I want") %>%
  
  #Add subway stations:
  addCircles(data = subway_stations, group = "See Subway Stations", radius = 20,
             color="red", opacity=1.0, fill = TRUE, fillColor = "red", fillOpacity=1.0,
             stroke=FALSE, label = stations_labels, labelOptions=labelOptions(textsize = "12px",
                                                    direction='top', offset=c(0,-10))) %>%
    # addCircleMarkers(data = subway_stations, label = stations_labels, group = "See Subway Stations", radius=7,color="red",stroke=FALSE,fillOpacity=0.5) %>%
  
  #Add subway Lines:  
  addPolylines(data = subway_lines, group="See Subway Lines",label = line_labels,
               labelOptions=labelOptions(textsize = "12px", direction='top',
                                         offset=c(0,-10)),
               color=ifelse(subway_lines$name=="4-5-6","#00933C",
                           ifelse(subway_lines$name=="Q","#FCCC0A","grey"))) %>%
  
  #2001-2005 Layer:
  addPolygons(data = NYC_Map2,group="2001-2005",
              color = "#444444",
              weight = 1,
              #smoothFactor = 0.2,
              opacity = 1.0,
              fillOpacity = 0.7,
              fillColor = ~palette(Y_2005),
              # change display properties while mouse cursor hovers on the polygon
              highlightOptions = highlightOptions(color="orange", 
                                                  weight=2, opacity=1, fillOpacity=1,
                                                  bringToFront = TRUE, sendToBack = TRUE),
              label = block_labels2005,
              # label display options:
              labelOptions = labelOptions(style = list("font-weight" = "normal", 
                                          padding = "3px 8px"),
                                          textsize = "12px",
                                          direction = "top",
                                          offset=c(0,-10))) %>%
  
  #2001-2010 Layer:
  addPolygons(data = NYC_Map2,group="2001-2010",
              color = "#444444",
              weight = 1,
              #smoothFactor = 0.2,
              opacity = 1.0,
              fillOpacity = 0.7,
              fillColor = ~palette(Y_2010),
              # change display properties while mouse cursor hovers on the polygon
              highlightOptions = highlightOptions(color="orange", 
                                                  weight=2, opacity=1, fillOpacity=1,
                                                  bringToFront = TRUE, sendToBack = TRUE),
              label = block_labels2010,
              # label display options:
              labelOptions = labelOptions(style = list("font-weight" = "normal", 
                                          padding = "3px 8px"),
                                          textsize = "12px",
                                          direction = "top",
                                          offset=c(0,-10))) %>%
  
  #2001-2015 Layer:
  addPolygons(data = NYC_Map2,group="2001-2015",
              color = "#444444",
              weight = 1,
              #smoothFactor = 0.2,
              opacity = 1.0,
              fillOpacity = 0.7,
              fillColor = ~palette(Y_2015),
              # change display properties while mouse cursor hovers on the polygon
              highlightOptions = highlightOptions(color="orange", 
                                                  weight=2, opacity=1, fillOpacity=1,
                                                  bringToFront = TRUE, sendToBack = TRUE),
              label = block_labels2015,
              # label display options:
              labelOptions = labelOptions(style = list("font-weight" = "normal", 
                                          padding = "3px 8px"),
                                          textsize = "12px",
                                          direction = "top",
                                          offset=c(0,-10))) %>%
  
  # legend 2001-2015
  leaflet::addLegend(pal = palette,
            values = NYC_Map2$Y_2015,
            labFormat = labelFormat(prefix="",suffix="%",between=", ",transform=function(x) 100*x),
            group = "2001-2015",
            opacity = 0.7,
            title = "Avg. increase in Bldg. Permits",
            position = "bottomright") %>%
  
  #legend for the Subway Lines:
  addLegend("topleft",
  colors =c("#FCCC0A","#00933C"),
  labels= c("2nd Ave Line", "Pre-existing 4th Avenue Line"),
  title= "Subway Lines",
  group="See Subway Lines",
  opacity = 0.7) %>%

  # Layers control
  addLayersControl(
    baseGroups = c("2001-2005","2001-2010","2001-2015"),
    overlayGroups = c("See Subway Lines", "See Subway Stations"),
    position = "topleft",
    options = layersControlOptions(collapsed = FALSE))

#myNYCmap <- myNYCmap %>% hideGroup("Building Permits 2001-2015")
# myNYCmap <- myNYCmap %>% hideGroup("See Subway Stations")
# myNYCmap <- myNYCmap %>% hideGroup("See Subway Lines")

myNYCmap  # Display the map

# myNYCmap %>% hideGroup("See Subway Lines")
# myNYCmap %>% hideGroup("See Subway Stations")
# myNYCmap %>% hideGroup("Building Permits 2001-2015")

htmlwidgets::saveWidget(myNYCmap, 'Interactive_Bldg_Permits_Map-Html Widget.html', selfcontained = TRUE)

```
<br>

#####**Our Conclusions:** 

After [a long Journey exploring the data](EDA-Building_Permits.html "Our EDA Journey") we came to the conclusion that, it seems that the number of building permits filed, is related to the distance in blocks from the phase 1 of the 2nd Ave Line subway project.

In the interactive map above, if we click to select the complete 2001-2015 period, we can clearly see that the average percentage of increase in filed building permits is higher for buildings within 1 block of distance to the Line and lower for buildings further away from the Line.

As an interesting observation (after selecting the complete 2001-2015 period), if we look with attention we can see that, while for buildings located to left, the average increase in building permits decreases with the distance (in blocks) from the 2nd Ave Line, for buildings to right, we can see an initial decrease from one block distance to 2 blocks distance, but it slightly increases back for buildings located within 3-4 blocks to the right of the line. During [our exploratory data analysis](EDA-Building_Permits.html "Our EDA Journey"), we concluded that this seems to be caused due to the fact that this last category includes buildings that are right on the edge of the East River, which attract a higher number of projects, due to the water views. 

See: [Our Exploratory Data Analysis (EDA) Journey](EDA-Building_Permits.html "Our EDA Journey")